A Convolutional Neural Network for Outsole Recognition\tnoteref{t1}

### 2020-04-27 paper/paper.Rnw
Updates to try to debug compiler

Collection and use of footwear evidence varies heavily across jurisdictions within the United States, and, though it is more common for such analyses to be used in Europe, the U.K., and Israel, many of the same fundamental issues remain. It is often difficult to recover footwear evidence with sufficient detail to identify individualizing characteristics, either because of degradation or because the medium did not record the individualizing characteristics in sufficient detail to examine. It is also common for prints to be only partially recorded or recovered, which further compounds the difficulty of performing these types of analysis. As a result, the majority of footwear examinations in the United States reach conclusions regarding the correspondence of class characteristics rather than of individualizing characteristics \citep{smith_2011}. This tendency is also reflected in the SWGTREAD range of conclusions scale, which has two separate categories to indicate degrees of class characteristic based association out of a range of 7 conclusions\citep{SWGTREAD2013}. Currently, the guidelines also caution against any statistical assessment because accurate data and models to support statistical calculations do not exist. However, if it were possible to collect a sufficiently large quantity of footwear impression data from a local population, it would be helpful to be able to calculate the frequency of shoes in the population that have similar class characteristics to a given shoe or print in order to characterize the determinative value of that particular piece of evidence. Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns on the class characteristic level.

@@

\paragraph{Transfer Learning}
Neural networks are composed of sets of layers; the early layers, which are called the model base, contain feature detectors, and the final layers make up the classifier, or model head, which connects meaningful features to classification labels (see \autoref{fig:VGG16-structure} for an example of the structure of a CNN). Each of these layers is composed of sets of weights and connections that are optimized during training. A relatively simple convolutional neural network which processes image data may contain more than 14 million parameters in the model base and an additional 120 million parameters in the model head; to train a model of this nature requires millions of labeled images and a significant amount of computational power. Assembling sufficient data to train a network from scratch is a gargantuan task; the process of \emph{transfer learning} is a natural solution to this common problem. Transfer learning leverages the modularity of neural networks, that is, that the model base that detects features can be separated from the classifier which produces predictions. Transfer learning approaches use the base of a model trained on more general image data and train a new model head or classifier using a smaller set of domain-specific labeled images\citep{oquabLearningTransferringMidlevel2014}. This allows CNNs to be applied to smaller datasets of several thousand images, reduces the amount of computational time required to fit the model, and provides boosted performance compared to training a new model from scratch \citep{yosinskiHowTransferableAre2014}. Transfer learning has been successfully applied in automatic classification of medical images \citep{shinDeepConvolutionalNeural2016} as well as in various applications of shoe forensics \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017}.

>>>>

Collection and use of footwear evidence varies heavily across jurisdictions within the United States, and, though it is more common for such analyses to be used in Europe, the U.K., and Israel, many of the same fundamental issues remain. It is often difficult to recover footwear evidence with sufficient detail to identify individualizing characteristics, either because of degradation or because the medium did not record the individualizing characteristics in sufficient detail to examine. It is also common for prints to be only partially recorded or recovered, which further compounds the difficulty of performing these types of analysis. As a result, the majority of footwear examinations in the United States reach conclusions regarding the correspondence of class characteristics rather than of individualizing characteristics \citep{smith_2011}. This tendency is also reflected in the SWGTREAD range of conclusions scale, which has two separate categories to indicate degrees of class characteristic based association out of a range of 7 conclusions \citep{SWGTREAD2013}. Currently, the guidelines also caution against any statistical assessment because accurate data and models to support statistical calculations do not exist. However, if it were possible to collect a sufficiently large quantity of footwear impression data from a local population, it would be helpful to be able to calculate the frequency of shoes in the population that have similar class characteristics to a given shoe or print in order to characterize the determinative value of that particular piece of evidence. Note that this question is distinct from make and model determination since many shoes of different makes and models can still have similar tread patterns on the class characteristic level.

@@

\paragraph{Transfer Learning}
Neural networks are composed of sets of layers; the early layers, which are called the model base, contain feature detectors, and the final layers make up the classifier, or model head, which connects meaningful features to classification labels (see \autoref{fig:VGG16-structure} for an example of the structure of a CNN). Each of these layers is composed of sets of weights and connections that are optimized during training. A relatively simple convolutional neural network which processes image data may contain more than 14 million parameters in the model base and an additional 120 million parameters in the model head; to train a model of this nature requires millions of labeled images and a significant amount of computational power. Assembling sufficient data to train a network from scratch is a gargantuan task; the process of \emph{transfer learning} is a natural solution to this common problem. Transfer learning leverages the modularity of neural networks, that is, that the model base that detects features can be separated from the classifier which produces predictions. Transfer learning approaches use the base of a model trained on more general image data and train a new model head or classifier using a smaller set of domain-specific labeled images \citep{oquabLearningTransferringMidlevel2014}. This allows CNNs to be applied to smaller datasets of several thousand images, reduces the amount of computational time required to fit the model, and provides boosted performance compared to training a new model from scratch \citep{yosinskiHowTransferableAre2014}. Transfer learning has been successfully applied in automatic classification of medical images \citep{shinDeepConvolutionalNeural2016} as well as in various applications of shoe forensics \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017}.

### 2020-04-27 paper/paper.tex
Updates to try to debug compiler

{\centering \includegraphics[width=\maxwidth]{/home/susan/Projects/Work/2019-CoNNORFSI/paper/images/labeled_filters}

@@

{\centering \includegraphics[width=.95\textwidth]{/home/susan/Projects/Work/2019-CoNNORFSI/paper/images/heatmaps/processed/heatmap-a7-Ugg-grey}

@@

\caption{A section of an UGG shoe containing the logo. The model correctly identifies triangles and text (and references the correct portion of the image in both cases). The model identifies a circle as well; while it could be argued that the triangles form an implicit circle, the heatmap demonstrates that the circular portion of the image is the ``G", which would not be a closed figure labeled as a circle. The model's prediction of the presence of a circle is understandable, but incorrect.\label{fig:ugg-heatmap}}\label{fig:current-heatmaps1}
\end{figure}

@@

{\centering \includegraphics[width=.95\textwidth]{/home/susan/Projects/Work/2019-CoNNORFSI/paper/images/heatmaps/processed/heatmap-text-seychelles}

@@

\caption{In this image of a Seychelles shoe, the model identifies text and circle. The prediction of text is correct; the prediction of circle is not---there are very round shapes, but they do not form a closed figure that could be identified as a circle.\label{fig:seychelles-heatmap}}\label{fig:current-heatmaps2}
\end{figure}

@@

{\centering \includegraphics[width=.95\textwidth]{/home/susan/Projects/Work/2019-CoNNORFSI/paper/images/heatmaps/processed/heatmap-04-circle_quad-2-dc-vestrey-black-camo_product_9065558_color_12784}

@@

\caption{This image from a DC shoe contains repeating instances of a shape that appears to be a combination of a circle and quadrilateral, but does not technically fall into any of the nine categories used in our labeling scheme. The model assigns high probabilities of circle and quadrilateral and indicates that the proper image features are being used for this classification. The prediction of polygon is incorrect but reasonable considering that many small hexagons are indistinguishable from circles.\label{fig:dc-heatmap}}\label{fig:current-heatmaps3}
\end{figure}

>>>>

{\centering \includegraphics[width=\maxwidth]{images/labeled_filters}

@@

{\centering \includegraphics[width=.95\textwidth]{images/heatmaps/processed/heatmap-a7-Ugg-grey}

@@

\caption{A section of an UGG shoe containing the logo. The model correctly identifies triangles and text (and references the correct portion of the image in both cases). The model identifies a circle as well; while it could be argued that the triangles form an implicit circle, the heatmap demonstrates that the circular portion of the image is the ``G", which would not be a closed figure labeled as a circle. The model's prediction of the presence of a circle is understandable, but incorrect.\label{fig:ugg-heatmap}}\label{fig:current-heatmaps}
\end{figure}

@@

{\centering \includegraphics[width=.95\textwidth]{images/heatmaps/processed/heatmap-text-seychelles}

@@

\caption{In this image of a Seychelles shoe, the model identifies text and circle. The prediction of text is correct; the prediction of circle is not---there are very round shapes, but they do not form a closed figure that could be identified as a circle.\label{fig:seychelles-heatmap}}\label{fig:current-heatmaps}
\end{figure}

@@

{\centering \includegraphics[width=.95\textwidth]{images/heatmaps/processed/heatmap-04-circle_quad-2-dc-vestrey-black-camo_product_9065558_color_12784}

@@

\caption{This image from a DC shoe contains repeating instances of a shape that appears to be a combination of a circle and quadrilateral, but does not technically fall into any of the nine categories used in our labeling scheme. The model assigns high probabilities of circle and quadrilateral and indicates that the proper image features are being used for this classification. The prediction of polygon is incorrect but reasonable considering that many small hexagons are indistinguishable from circles.\label{fig:dc-heatmap}}\label{fig:current-heatmaps}
\end{figure}
